Self-sustained carbon monoxide oxidation oscillations on size-selected platinum nanoparticles at atmospheric pressure

### 2012-11-30 paper.tex
Now we are close to being able to re-submit

Already the early theretical descriptions of the possibility of oscillationary beheaviour even very simple catalytic systems (eg \cite{LAGOS1979,IVANOV1980,TAKOUDIS1981,SONG1991}) noticed that even though oscillation can theoretically occur in many catalytic systems, the details are extremely specific to the given system, and it can thus be diffuctult to predict how and if a given system will oscillate.

Oscillations in catalytic systems can be quite complex and the parameter space can be very large\cite{Ferino1999}. In the system investigated in this paper, oscillations could be dependant on morphologic parameres of the nanoparticels such as size, shape and exposed facets. Also the oxidation state and possible carbon pollution on the surface could have impact on the system. More macroscopic parameters such as the geometry of the reactor, the temperature distribution and the distribution of the particles throughout the reactor will influcence the systems' ability to show kinetic oscillations.

@@

In this study Pt nanoparticles were investigated which, compared to a single crystal or a thin film, is a very rough surface. It is to be expected that the particles will initially favor the reduced state. As the reaction runs the gas composition towards the outlet of the reactor will become more oxidizing increasing the oxidation rate of the particles (illustrated in Figure~\ref{fgr:full_oscillation}). Gradually, as the particles toward the outlet oxidizes the turn-over will increase and the general CO concentration will decrease promoting more and more particles to oxidize which will gradually be seen in the QMS as an increase in CO$_2$ and a decrease in CO. As the CO concentration decreases the light-off temperature will decrease hence approaching the constant temperature of the sample resulting in a steep increase in reactivity. It is important to note that the fraction of oxidized particles needed to achieve full conversion is not known. We are currently investigating new methods for gaining insight into the nature of the nanoparticles in the reactor but the low coverage and the design of the reactor !!! and the fact that the lid can only be removed by mechanically breaking the reactor!!! makes this a challenging task. If not all particles are oxidized at light-off they will now oxidize much faster due to the low CO-concentration. The further increased rate will not be visible in the QMS since the reactor is already in full conversion. As most of the particles are now oxidized they will become more and more roughened and gradually return to the reduced state. Fewer and fewer oxidized particles will be responsible for the activity increasing the roughing rate of the remaining active particles. Just before the overall activity drops only a fraction of the particles participates in the reaction and the deactivation will happen very suddenly. Once these potentially few particles eventually becomes sufficiently rough they will reduce and lose activity hence completing the oscillation cycle. It should be noted that our group has recently measured turnover frequencies of CO oxidation on Pt of almost $10^{5}$s$^{-1}$ per site corresponding to as little as $\sim$50000 Pt nanoparticles contributing to the reactivity just before the end of the high-activity cycle. This corresponds to an area of $\sim$0.5 ppm of the reactor area suggesting that in situ studies become challenging as such a minority of the reactor volume could account for all the reactivity. Attempts to see any thermal signal from such small areas has so far been in vain.

The oscillation period increases with time and becomes increasingly less stable as shown in Figure~\ref{fgr:extracts}. This phenomena is consistent across all measured samples. The large scatter in oscillation period can be attributed to the smaller oscillations in between on/off cycles. Occasionally, these small amplitude oscillations will trigger a full switch thus introducing more short-period oscillations despite the trend of increasingly slower oscillation period. A possible explanation of the phenomeonon could be sintering of the particles during the oxidation and reductions cycles. However, since the average conversion integrated over full oscillation periods is constant we do not expect sintering to be the cause since an average loss in activity of the sample would then be expected.

>>>>

Already the early theoretical descriptions of the possibility of oscillationary behavior of even very simple catalytic systems (eg \cite{LAGOS1979,IVANOV1980,TAKOUDIS1981,SONG1991}) noticed that even though oscillation can theoretically occur in many catalytic systems, the details are extremely specific to the given system, and it can thus be difficult to predict how and if a given system will oscillate.

Oscillations in catalytic systems can be quite complex and the parameter space can be very large\cite{Ferino1999}. In the system investigated in this paper, oscillations could be dependant on morphologic parameters of the nanoparticels such as size, shape and exposed facets. Also the oxidation state and possible carbon pollution on the surface could have impact on the system. More macroscopic parameters such as the geometry of the reactor, the temperature distribution and the distribution of the particles throughout the reactor will also influcence the systems ability to show kinetic oscillations.

@@

In this study Pt nanoparticles were investigated which, compared to a single crystal or a thin film, is a very rough surface. It is to be expected that the particles will initially favor the reduced state. As the reaction runs the gas composition towards the outlet of the reactor will become more oxidizing increasing the oxidation rate of the particles (illustrated in Figure~\ref{fgr:full_oscillation}). Gradually, as the particles toward the outlet oxidizes the turn-over will increase and the general CO concentration will decrease promoting more and more particles to oxidize which will gradually be seen in the QMS as an increase in CO$_2$ and a decrease in CO. As the CO concentration decreases the light-off temperature will decrease hence approaching the constant temperature of the sample resulting in a steep increase in reactivity. It is important to note that the fraction of oxidized particles needed to achieve full conversion is not known. We are currently investigating new methods for gaining insight into the nature of the nanoparticles in the reactor but the low coverage and the design of the reactor and the fact that the lid can only be removed by mechanically breaking the reactor makes this a challenging task. If not all particles are oxidized at light-off they will now oxidize much faster due to the low CO-concentration. The further increased rate will not be visible in the QMS since the reactor is already in full conversion. As most of the particles are now oxidized they will become more and more roughened and gradually return to the reduced state. Fewer and fewer oxidized particles will be responsible for the activity increasing the roughing rate of the remaining active particles. Just before the overall activity drops only a fraction of the particles participates in the reaction and the deactivation will happen very suddenly. Once these potentially few particles eventually becomes sufficiently rough they will reduce and lose activity hence completing the oscillation cycle. It should be noted that our group has recently measured turnover frequencies of CO oxidation on Pt of almost $10^{5}$s$^{-1}$ per site corresponding to as little as $\sim$50000 Pt nanoparticles contributing to the reactivity just before the end of the high-activity cycle. This corresponds to an area of $\sim$0.5 ppm of the reactor area suggesting that in situ studies become challenging as such a minority of the reactor volume could account for all the reactivity. Attempts to see any thermal signal from such small areas has so far been in vain, which would point towards a constant temperature even on the microscopic scale, however the experimental data are so far not good enough to make final conclusions.

The oscillation period increases with time and becomes increasingly less stable as shown in Figure~\ref{fgr:extracts}. This phenomena is consistent across all measured samples. The large scatter in oscillation period can be attributed to the smaller oscillations in between on/off cycles. Occasionally, these small amplitude oscillations will trigger a full switch thus introducing more short-period oscillations despite the trend of increasingly slower oscillation period. A possible explanation of the phenomeonon could be sintering of the particles during the oxidation and reductions cycles. However, since the average conversion integrated over full oscillation periods is constant we do not expect sintering to be the cause since an average loss in activity of the sample would then be expected. Another possible mechanism behind the increasing oscillation period, could be morphology changes of the nano particles. However, the fact that the oscillations show no clear size dependence in the range of 3 to 9\,nm suggests that the oscillations are not very dependent on the exact particle morphology.
